Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-08-02 analysis/paper/methods.Rmd
added Tian et al 2016 citation

@@

For each hotspot, we identified a set of local expression traits and a
set of nonlocal expression traits. All traits demonstrated LOD peaks that met
the genome-wide significance threshold, 7.18 [@keller2018genetic].
The nonlocal expression traits arose from genes
on other chromosomes, while the local traits reflected expression levels for
genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
the hotspot center.

@@

A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.

For each hotspot and each local trait - nonlocal trait pair, we also performed a
mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait. Each mediation analysis involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].

>>>>

We used the publicly available data from @keller2018genetic to demonstrate our strategy. @keller2018genetic
measured genome-wide gene expression levels with RNA sequencing of pancreatic islets in 378 Diversity Outbred mice. From the RNA sequencing counts, they obtained normalized gene expression levels for 21,771 transcripts, which they shared publicly via DataDryad ().

@@

For each hotspot, we identified a set of local expression traits and a
set of nonlocal expression traits. We focused on those local 
traits that strongly associated with QTL within the hotspot. We set the univariate LOD threshold for local traits at 15. We limited consideration to those local traits for which the middle of the gene lay within the hotspot interval. 
All nonlocal traits demonstrated univariate LOD peaks that met
the genome-wide significance threshold, 7.18 [@keller2018genetic]. The nonlocal expression traits arose from genes
on other chromosomes, while the local traits reflected expression levels for
genes on the same chromosome as the hotspot . We ignored transcripts for genes on the hotspot chromosome but outside the hotspot interval.

@@

A mediation analysis in systems genetics begins with an identified QTL-expression trait 
association [@chick2016defining]. In our case, we choose a QTL that affects a nonlocal trait, then ask if any of 
the nearby local traits lie in the causal pathway from QTL to nonlocal trait.

involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
